\section{Properties of Infinite Series}

\begin{exercise}
  Proving the Alternating Series Test (Theorem 2.7.7) amounts to showing that the sequence of partial sums
  $$
  s_{n}=a_{1}-a_{2}+a_{3}-\cdots \pm a_{n}
  $$
  converges. (The opening example in Section $2.1$ includes a typical illustration of $\left(s_{n}\right)$.) Different characterizations of completeness lead to different proofs.

  \enum{
  \item Prove the Alternating Series Test by showing that $\left(s_{n}\right)$ is a Cauchy sequence.
  \item Supply another proof for this result using the Nested Interval Property (Theorem 1.4.1).
  \item Consider the subsequences $\left(s_{2 n}\right)$ and $\left(s_{2 n+1}\right)$, and show how the Monotone Convergence Theorem leads to a third proof for the Alternating Series Test.
  }
\end{exercise}

\begin{solution}
  \enum{
  \item Let $N \in \mathbf N$ be even and let $n \ge N$. because the series is alternating we have
    $$
    s_N \le s_n \le s_{N+1}
    $$
    Obviously $|s_{N+1} - s_N| = |a_N|$ can be made as small as we like by increasing $N$, setting $N$ large enough to make $|a_N| < \epsilon/2$ gives
    $$
    |s_m - s_n| \le |s_m - s_N| + |s_N - s_n| < \epsilon/2 + \epsilon/2 = \epsilon
    $$
    Which shows $(s_n)$ is cauchy, and hence converges by the Cauchy Criterion.
  \item Let $I_1$ be the interval $[a_1 - a_2, a_1]$ and in general $I_n = [a_n - a_{n+1}, a_n]$, since $(a_n)$ is decreasing we have $I_{n+1} \subseteq I_n$. Applying the nested interval property gives
    $$\bigcap_{n=1}^\infty I_n \ne \emptyset$$
    Let $x \in \bigcap_{n=1}^\infty I_n$, since $a_n \in I_n$ and $x \in I_n$ the distance $|a_n - x|$ must be less then the length $|I_n|$. and since the length goes to zero $|a_n - x|$ can be made less then any $\epsilon$.
  \item If we can show $\lim s_{2n} = \lim s_{2n+1} = s$ that will imply $\lim s_n = s$ since each $s_n$ is either in $(s_{2n+1})$ or in $(s_{2n})$ as $n$ is must be even or odd.

    We have $s_{2n+1} \le a_1$ since
    $$
    s_{2n+1} = a_1 - (a_2 - a_3) - \dots - (a_{2n} - a_{2n+1}) \le a_1
    $$
    Thus $s_{2n+1} \to s$ by the Monotone Convergence Theorem, to show $(s_{2n}) \to s$ notice $s_{2n} = s_{2n+1} - a_{2n+1}$ with $(a_{2n+1}) \to 0$ meaning we can use the triangle inequality
    $$
    |s_{2n} - s| \le \underbrace{|s_{2n} - s_{2n+1}|}_{a_{2n+1}} + |s_{2n+1} - s| < \epsilon/2 + \epsilon/2 < \epsilon
    $$
    Thus $(s_{2n}) \to s$ aswell finally implying $(s_n) \to s$.

    \textbf{Summary:} Partition the alternating series into two subsequences of partial sums, then use MCT to show they both converge to the same limit.
  }
\end{solution}

\begin{exercise}
  Decide whether each of the following series converges or diverges:
  \enum{
  \item $\sum_{n=1}^{\infty} \frac{1}{2^{n}+n}$
  \item $\sum_{n=1}^{\infty} \frac{\sin (n)}{n^{2}}$
  \item $1-\frac{3}{4}+\frac{4}{6}-\frac{5}{8}+\frac{6}{10}-\frac{7}{12}+\cdots$
  \item $1+\frac{1}{2}-\frac{1}{3}+\frac{1}{4}+\frac{1}{5}-\frac{1}{6}+\frac{1}{7}+\frac{1}{8}-\frac{1}{9}+\cdots$
  \item $1-\frac{1}{2^{2}}+\frac{1}{3}-\frac{1}{4^{2}}+\frac{1}{5}-\frac{1}{6^{2}}+\frac{1}{7}-\frac{1}{8^{2}}+\cdots$
  }
\end{exercise}

\begin{solution}
  \enum{
  \item Converges by a comparison test with $\sum_{n=1}^\infty \frac{1}{2^n}$.
  \item Converges by a comparison test with $\sum_{n=1}^\infty \frac{1}{n^2}$.
  \item Diverges since $(n+1)/2n = 1/2 + 1/2n$ never gets smaller then $1/2$.
  \item Grouping terms gives
    $$
    \frac{1}{n} + \left(\frac{1}{n+1} - \frac{1}{n+2}\right) \ge \frac{1}{n}
    $$
    Which shows the subsequence $(s_{3n})$ diverges (via comparison test with the harmonic series) hence $(s_n)$ also diverge.
  \item Intuitively this should diverge since it is a mixture of $1/n$ (divergent) and $1/n^2$ (convergent). To make this rigorous examine the subsequence $(s_{2n})$
    $$
    s_{2n} = 1 - 1/2^2 + 1/3 - 1/4^2 + \dots + 1/(2n)^2
    $$
    Let $t_n = \sum_{k=1}^n \frac{1}{2k-1}$ and $v_n = \sum_{k=1}^n \frac{1}{(2k)^2}$ so $s_{2n} = t_n - v_n$.

    A comparison test with the harmonic series (after some manipulation) shows that $(t_n)$ diverges, and p-series tells us $(v_n)$ converges. Therefor their difference $s_{2n} = t_n - v_n$ must diverge, which implies $(s_n)$ diverges as desired.
  }
\end{solution}

\begin{exercise}
  \enum{
  \item Provide the details for the proof of the Comparison Test (Theorem 2.7.4) using the Cauchy Criterion for Series.
  \item Give another proof for the Comparison Test, this time using the Monotone Convergence Theorem.
  }
\end{exercise}

\begin{solution}
  Suppose $a_n, b_n \ge 0$, $a_n \le b_n$ and define $s_n = a_1 + \dots + a_n$, $t_n = b_1 + \dots + b_n$.

  \enum{
  \item We have $|a_{m+1} + \dots + a_n| \le |b_{m+1} + \dots + b_n| < \epsilon$ implying $\sum_{n=1}^\infty a_n$ converges by the cauchy criterion.
    The other direction is analogous, if $(s_n)$ diverges then $(t_n)$ must also diverge since $s_n \le t_n$.
  \item Since $(t_n) \to t$. This implies that $s_n$ is bounded, and since $s_n \le t_n$ implies $s_n \le t$ by the order limit theorem, we can use the monotone convergence theorem to conclude $(s_n)$ converges.
  }
\end{solution}

\begin{exercise}
  Give an example of each or explain why the request is impossible referencing the proper theorem(s).
  \enum{
  \item Two series $\sum x_{n}$ and $\sum y_{n}$ that both diverge but where $\sum x_{n} y_{n}$ converges.
  \item A convergent series $\sum x_{n}$ and a bounded sequence $\left(y_{n}\right)$ such that $\sum x_{n} y_{n}$ diverges.
  \item Two sequences $\left(x_{n}\right)$ and $\left(y_{n}\right)$ where $\sum x_{n}$ and $\sum\left(x_{n}+y_{n}\right)$ both converge but $\sum y_{n}$ diverges.
  \item A sequence $\left(x_{n}\right)$ satisfying $0 \leq x_{n} \leq 1 / n$ where $\sum(-1)^{n} x_{n}$ diverges.
  }
\end{exercise}

\begin{solution}
  \enum{
  \item $x_n = 1/n$ and $y_n = 1/n$ have their respective series diverge, but $\sum x_ny_n = \sum 1/n^2$ converges since it is a p-series with $p > 1$.
  \item Let $x_n = (-1)^n/n$ and $y_n = (-1)^n$. $\sum x_n$ converges but $\sum x_ny_n = \sum 1/n$ diverges.
  \item Impossible as the algebraic limit theorem for series implies $\sum(x_n+y_n) - \sum x_n = \sum y_n$ converges.
  \item The sequence
  \[x_n = \begin{cases}
      \frac{1}{n} \text{ if } n \text{ even} \\
      0 \text{ otherwise}
  \end{cases}\]
  }
   diverges for the same reason the harmonic series does.
\end{solution}

\begin{exercise}
  Prove the series $\sum_{n=1}^{\infty} 1 / n^{p}$ converges if and only if $p>1$. (Corollary 2.4.7)
\end{exercise}

\begin{solution}
  Eventually we have $1/n^p < 1/p^n$ for $p > 1$ (polynomial vs exponential) meaning we can use the comparison test to conclude $\sum_{n=1}^\infty 1/n^p$ converges if $p > 1$.

  Now suppose $p \le 1$, since $1/n^p \le 1/n$ a comparsion test with the harmonic series implies $p \le 1$ diverges.
\end{solution}

\begin{exercise}
  Let's say that a series subverges if the sequence of partial sums contains a subsequence that converges. Consider this (invented) definition for a moment, and then decide which of the following statements are valid propositions about subvergent series:
  \enum{
  \item If $\left(a_{n}\right)$ is bounded, then $\sum a_{n}$ subverges.
  \item All convergent series are subvergent.
  \item If $\sum\left|a_{n}\right|$ subverges, then $\sum a_{n}$ subverges as well.
  \item If $\sum a_{n}$ subverges, then $\left(a_{n}\right)$ has a convergent subsequence.
  }
\end{exercise}

\begin{solution}
  \enum{
    \item False, consider $a_n = 1$ then $s_n = n$ does not have a convergent subsequence.
    \item True, every subsequence converges to the same limit in fact.
    \item True, since $s_n = \sum_{k=1}^n |a_k|$ converges it is bounded $|s_n| \le M$, and since $t_n = \sum_{k=1}^n a_k$ is smaller $t_n \le s_n$ it is bounded $t_n \le M$ which by BW implies there exists a convergent subsequence $(t_{n_k})$.
    \item False, $a_n = (1, -1, 2, -2, \dots)$ has no convergent subsequence but the sum $s_n = \sum_{k=1}^n a_k$ has the subsequence $(s_{2n}) \to 0$.
  }
\end{solution}

\begin{exercise}
  \enum{
  \item Show that if $a_{n}>0$ and $\lim \left(n a_{n}\right)=l$ with $l \neq 0$, then the series $\sum a_{n}$ diverges.
  \item Assume $a_{n}>0$ and $\lim \left(n^{2} a_{n}\right)$ exists. Show that $\sum a_{n}$ converges.
  }
\end{exercise}

\begin{solution}
  Note: This is kind of like a wierd way to do a comparison with $1/n$ and $1/n^2$.
  \enum{
  \item If $\lim(na_n) = l \ne 0$ then $na_n \in (l-\epsilon, l+\epsilon)$, setting $\epsilon = l/2$ gives $na_n \in (l/2, 3l/2)$ implying $a_n > (l/2)(1/n)$. But if $a_n > (l/2)(1/n)$ then $\sum a_n$ diverges as it is a multiple of the harmonic series. (note that $a_n > 0$ ensures $l \ge 0$.)
  \item Letting $l = \lim(n^2a_n)$ we have $n^2a_n \in (l-\epsilon,l+\epsilon)$ setting $\epsilon = l$ gives $n^2a_n \in (0, 2l)$ implying $0 \le a_n \le 2l/n^2$ and so $\sum a_n$ converges by a comparsion test with $\sum 2l/n^2$.
  }
\end{solution}

\begin{exercise}
  Consider each of the following propositions. Provide short proofs for those that are true and counterexamples for any that are not.
  \enum{
  \item If $\sum a_{n}$ converges absolutely, then $\sum a_{n}^{2}$ also converges absolutely.
  \item If $\sum a_{n}$ converges and $\left(b_{n}\right)$ converges, then $\sum a_{n} b_{n}$ converges.
  \item If $\sum a_{n}$ converges conditionally, then $\sum n^{2} a_{n}$ diverges.
  }
\end{exercise}

\begin{solution}
  \enum{
  \item True since $(a_n) \to 0$ so eventually $a_n^2 \le |a_n|$ meaning $\sum a_n^2$ converges by a comparsion test with $\sum |a_n|$.
  \item False, let $a_n = (-1)^n/{\sqrt n}$ and $b_n = (-1)^n/{\sqrt n}$. $\sum a_n$ converges by the alternating series test, but $\sum a_nb_n = \sum 1/n$ diverges.
  \item True, suppose $(n^2a_n)$ converges, since $(n^2 a_n) \to 0$ we have $|n^2 a_n| < 1$ for $n > N$, implying $|a_n| < 1/n^2$. But if $|a_n| < 1/n^2$ then a comparsion test with $1/n^2$ implies $a_n$ converges absolutely, contradicting the assumption that $a_n$ converges conditionally. Therefor $\sum n^2a_n$ must diverge.
  }
\end{solution}

\begin{exercise}[Ratio Test]
  Given a series $\sum_{n=1}^{\infty} a_{n}$ with $a_{n} \neq 0$, the Ratio Test states that if $\left(a_{n}\right)$ satisfies
  $$
  \lim \left|\frac{a_{n+1}}{a_{n}}\right|=r<1
  $$
  then the series converges absolutely.

  \enum{
  \item Let $r^{\prime}$ satisfy $r<r^{\prime}<1$. Explain why there exists an $N$ such that $n \geq N$ implies $\left|a_{n+1}\right| \leq\left|a_{n}\right| r^{\prime}$.
  \item Why does $\left|a_{N}\right| \sum\left(r^{\prime}\right)^{n}$ converge?
  \item Now, show that $\sum\left|a_{n}\right|$ converges, and conclude that $\sum a_{n}$ converges.
  }
\end{exercise}

\begin{solution}
  \enum{
  \item We are given
    $$
    \left|\frac{a_{n+1}}{a_n} - r\right| < \epsilon
    $$
    Since $1 > r' > r$ we can set $\epsilon = r' - r$ meaning the neighborhood
    $$
    \frac{a_{n+1}}{a_n} \in (r-\epsilon, r+\epsilon) = (2r-r', r')
    $$
    Is all less then $r'$ meaning
    $$
    \left|\frac{a_{n+1}}{a_n}\right| \le r' \implies |a_{n+1}| \le r'|a_n|
    $$
  \item Let $N$ be large enough that for $n > N$ we have $|a_{n+1}| \le |a_n|r'$.
    Applying this multiple times gives $|a_n| \le (r')^{n-N} |a_N|$ which gives
    $$
    |a_N| + |a_{N+1}| + \dots + |a_n| \le |a_N| + r'|a_N| + \dots + (r')^{n-N}|a_N|
    $$
    Factoring out $|a_N|$ and writing with sums gives
    $$
    \sum_{k=N}^n |a_k| \le |a_N|\sum_{k=0}^{n-1} (r')^k
    $$
    Which converges as $n \to \infty$ since $|r'| < 1$ and $|a_N|$ is constant. Implying $\sum_{k=N}^n |a_k|$ converges and thus $\sum_{k=1}^n |a_k|$ also converges since we only omitted finitely many terms.
  \item See (b)
  }
\end{solution}

\begin{exercise}[Infinite Products]
  Review Exercise 2.4.10 about infinite products and then answer the following questions:
  \enum{
  \item Does $\frac{2}{1} \cdot \frac{3}{2} \cdot \frac{5}{4} \cdot \frac{9}{8} \cdot \frac{17}{16} \cdots$ converge?
  \item The infinite product $\frac{1}{2} \cdot \frac{3}{4} \cdot \frac{5}{6} \cdot \frac{7}{8} \cdot \frac{9}{10} \cdots$ certainly converges. (Why?) Does it converge to zero?
  \item In 1655, John Wallis famously derived the formula
    $$
    \left(\frac{2 \cdot 2}{1 \cdot 3}\right)\left(\frac{4 \cdot 4}{3 \cdot 5}\right)\left(\frac{6 \cdot 6}{5 \cdot 7}\right)\left(\frac{8 \cdot 8}{7 \cdot 9}\right) \cdots=\frac{\pi}{2}
    $$
    Show that the left side of this identity at least converges to something. (A complete proof of this result is taken up in Section 8.3.)
  }
\end{exercise}

\begin{solution}
  \enum{
  \item Rewriting the terms as $a_n = (1 + 1/n)$ and using the result from 2.4.10 implies the product diverges since $\sum 1/n$ diverges.
  \item Converges by the monotone convergence theorem since the partial products are decreasing and greater then zero. To show that the product converges to zero, the key insight is to rewrite each term \(a_n = (2n -1) / (2n) = 1 / (2n/(2n - 1)) = 1/b_n\), where \(b_n = 2n / (2n -1)\). Then the partial products
\[
    p_n = \prod^n_{i = 1} a_n = \frac{1}{\prod^n_{i = 1} b_n}
\]
But
\[\prod^n_{i=1}b_n = \prod^n_{i=1} \left(1 + \frac{1}{2n - 1}\right)\]
and since \(\sum 1/(2n-1)\) diverges by comparison against a multiple of the harmonic series, \(\prod b_n\) diverges. Thus, to show there exists some \(N\) so that \(n \geq N\) implies \(p_n < \epsilon\), simply take \(N\) large enough so that \(\prod^n_{i = 1} b_n > 1/\epsilon\).

  \item In compontent form we have
    $$
    a_n
    = \frac{(2n)^2}{(2n-1)(2n+1)}
    = \frac{(2n)^2}{(2n)^2 - 1}
    = 1 + \frac{(2n)^2 - ((2n)^2 - 1)}{(2n)^2 - 1}
    = 1 + \frac{1}{(2n)^2 - 1}
    $$
    And since $\sum 1/((2n)^2 - 1)$ converges by a comparison test with $1/n^2$, exercise 2.4.10 implies
    $$\prod_{n=1}^\infty \left(1 + \frac{1}{(2n)^2 - 1}\right)$$
    also converges.
  }
\end{solution}

\begin{exercise}
  Find examples of two series $\sum a_{n}$ and $\sum b_{n}$ both of which diverge but for which $\sum \min \left\{a_{n}, b_{n}\right\}$ converges. To make it more challenging, produce examples where $\left(a_{n}\right)$ and $\left(b_{n}\right)$ are strictly positive and decreasing.
\end{exercise}

\begin{solution}
  Let $m_n = \min\{a_n, b_n\}$.
  Clearly $(m_n)$ must take an infinite amount of $(a_n)$ and $(b_n)$ terms, as otherwise removing the finite terms would imply one of $\sum a_n$ or $\sum b_n$ converged.

    The key insight is that as long as \(a_n, b_n > 0\), we can simply repeat terms in one sequence (while letting \(m_n\) be governed by the other sequence) for as long as we want - say, until we have enough terms to e.g. sum to 1. Then we can switch the roles of the sequences.
    To start with the construction, take some converging series with all terms positive - say, \(m_n = 1/2^n\).
    Then, define the first few terms of \(m_n\), \(a_n\), and \(b_n\) as:

\[
\begin{array}{|c|c|c|c|c|c|c|c|c|c|c|}
    \hline
n   &1 &2 &3 &[4, 4 + 8 = 12)& [12, 12 + 2^{11}) & [12 + 2^{11}, 12 + 2^{11} + 2^{12 + 2^{11}})\\ \hline
m_n &1/2 &1/4 & 1/8 & 1/2^n & 1/2^n & 1/2^n \\ \hline
\min\{a_n, b_n\} &a_n &b_n & b_n & a_n & b_n & a_n\\ \hline
a_n &1/2 & 1/2 & 1/2 & 1/2^n & 1/2^{11} & 1/2^n\\ \hline
b_n &1 & 1/4 & 1/8 & 1/8 & 1/2^n & 1/2^{12 + 2^{11}}\\ \hline
\end{array}
\]

Specifically, \(\min\{a_n, b_n\}\) will alternate between following \(a_n\) and \(b_n\). Every ``block'' of the sequence (examples: each of the last three columns) that isn't being followed by \(\min\{a_n, b_n\}\) sums to 1.
Since each block is finite, \(\min\{a_n, b_n\}\) will alternate between \(a_n\) and \(b_n\) infinitely, and thus both \(\sum a_n\) and \(\sum b_n\) will diverge.

For the sake of completeness, \(a_n\) and \(b_n\) are defined more formally below.
Let \(k_1 = 1\), \(k_n = k_{n-1} + 2^{k_{n-1} - 1}\). Then
\[
    a_n = \begin{cases}
        1/2^n & \text{ if } k_{2p - 1} \leq n < k_{2p} \\
        1/2^{k_{2p} - 1} &\text{ if } k_{2p} \leq n < k_{2p + 1}
    \end{cases}
\text{ and }
    b_n = \begin{cases}
        1/2^n & \text{ if } k_{2p} \leq n < k_{2p + 1} \\
        1/2^{k_{2p - 1} - 1} &\text{ if } k_{2p - 1} \leq n < k_{2p}
    \end{cases}
\]


% Preliminary results/requirements on a_n and b_n:
%   Let $m_n = \min\{a_n, b_n\}$.
%   Clearly $(m_n)$ must take an infinite amount of $(a_n)$ and $(b_n)$ terms, as otherwise removing the finite terms would imply one of $\sum a_n$ or $\sum b_n$ converged.

%   Because $(m_n) \to 0$ and because $(m_n)$ contains an infinite amount of $(a_n)$ terms we can find $N$ large enough that $a_N < \epsilon$, because $(a_n)$ is decreasing this implies every $n \ge N$ has $a_n < \epsilon$, which means $(a_n) \to 0$. The same logic shows $(b_n) \to 0$.

%   If we forget about the decreasing requirement we can make $a_n$ alternate between $1/n$ and $1/n^2$ and have $b_n$ do the same, then $m_n = 1/n^2$ converges.

%   A similar example won't work here, since if $\sum a_{2n-1}$ converges then $a_{2n} < a_{2n-1}$ implies $\sum a_{2n}$ converges by the comparison test, combining these shows $\sum a_n$ converges. Therefor $(m_n)$ can't alternate terms, and we need to be more clever.

%   In fact, If $\sum a_{n_k}$ converges then every shifted sum $\sum a_{n_k+p}$ will converge by a similar argument as above. Meaning $\left(a_{n_k}\right)$ can't be periodic since that would make $(a_n)$ converge.

  % TODO: Same exercise for functions? add appendix with my own exercises
\end{solution}

\begin{exercise}[Summation-by-parts]
  Let $\left(x_{n}\right)$ and $\left(y_{n}\right)$ be sequences, let $s_{n}=x_{1}+x_{2}+\cdots+x_{n}$ and set $s_{0}=0 .$ Use the observation that $x_{j}=s_{j}-s_{j-1}$ to verify the formula
  $$
  \sum_{j=m}^{n} x_{j} y_{j}=s_{n} y_{n+1}-s_{m-1} y_{m}+\sum_{j=m}^{n} s_{j}\left(y_{j}-y_{j+1}\right)
  $$
\end{exercise}

\begin{solution}
  Since $x_j = s_j - s_{j-1}$ we can rewrite the sum as
  $$
  \sum_{j=m}^n x_jy_j
  = \sum_{j=m}^n y_j(s_j - s_{j-1})
  = s_ny_{n+1} - s_my_{m-1} + \sum_{j=m}^n s_j(y_j - y_{j+1})
  $$
  In the last part we combine each $s_jy_j$ term with the $-s_{j}y_{j+1}$ term which is next in the sum, then we add some correction terms for the start and ending points.

  Note the symmetry here, we can turn a sum $\sum y_j(s_j-s_{j-1})$ into a sum $\sum s_j(y_j-y_{j+1})$ (at the cost of some correction terms). This is a useful pattern to keep in mind.
\end{solution}

\begin{exercise}[Abel's Test]
  Abel's Test for convergence states that if the series $\sum_{k=1}^{\infty} x_{k}$ converges, and if $\left(y_{k}\right)$ is a sequence satisfying
  $$
  y_{1} \geq y_{2} \geq y_{3} \geq \cdots \geq 0
  $$
  then the series $\sum_{k=1}^{\infty} x_{k} y_{k}$ converges.
  \enum{
  \item Use Exercise $2.7 .12$ to show that
    $$
    \sum_{k=1}^{n} x_{k} y_{k}=s_{n} y_{n+1}+\sum_{k=1}^{n} s_{k}\left(y_{k}-y_{k+1}\right)
    $$
    where $s_{n}=x_{1}+x_{2}+\cdots+x_{n}$.
  \item Use the Comparison Test to argue that $\sum_{k=1}^{\infty} s_{k}\left(y_{k}-y_{k+1}\right)$ converges absolutely, and show how this leads directly to a proof of Abel's Test.
  }
\end{exercise}

\begin{solution}
  \enum{
  \item Exercise 2.7.12 combined with $s_0=0$ gives
    $$
    \sum_{k=1}^n x_ky_k = s_ny_{n+1} + \sum_{k=1}^n s_k(y_k - y_{k+1})
    $$
    as desired.
  \item $s_ny_{n+1}$ clearly converges since $y_{n+1}$ is ``eventually constant'', so we must only show the right hand side converges.

    We will show absolute convergence, note $y_k - y_{k+1} \ge 0$ and so
    $$
    \sum_{k=1}^n |s_k|(y_k - y_{k+1}) \ge 0
    $$
    Bounding $|s_n| \le M$ gives
    $$
    \sum_{k=1}^n |s_k|(y_k - y_{k+1}) \le M \sum_{k=1}^n (y_k-y_{k+1})
    $$
    Since $\sum_{k=1}^n (y_k - y_{k+1}) = y_1 - y_{n+1}$ is telescoping we can write
    $$
    \sum_{k=1}^n |s_k|(y_k - y_{k+1}) \le M(y_1 - y_{n+1}) \le My_1
    $$
    Implying $\sum_{k=1}^\infty |s_k|(y_k - y_{k+1})$ converges since it is bounded and increasing.
    And since the series converges absolutely so does the original $\sum_{k=1}^\infty s_k(y_k - y_{k+1})$.

    \textbf{Summary:} Bound $|s_k| \le M$ and use the fact that $\sum(y_k - y_{k+1})$ is telescoping.
  }
\end{solution}

\begin{exercise}[Dirichlet's Test]
  Dirichlet's Test for convergence states that if the partial sums of $\sum_{k=1}^{\infty} x_{k}$ are bounded (but not necessarily convergent), and if $\left(y_{k}\right)$ is a sequence satisfying $y_{1} \geq y_{2} \geq y_{3} \geq \cdots \geq 0$ with $\lim y_{k}=0$, then the series $\sum_{k=1}^{\infty} x_{k} y_{k}$ converges.
  \enum{
  \item Point out how the hypothesis of Dirichlet's Test differs from that of Abel's Test in Exercise 2.7.13, but show that essentially the same strategy can be used to provide a proof.
  \item Show how the Alternating Series Test (Theorem 2.7.7) can be derived as a special case of Dirichlet's Test.
  }
\end{exercise}

\begin{solution}
  \enum{
  \item
    Abel's test gets it's convergence from $\sum x_n$ converging, while Dirichlet's test gets its convergence from $(y_n) \to 0$.
    Expanding on that, the proof that the $\sum_{k=1}^n s_k(y_k-y_{k-1})$ term converges is the same in Abel and Dirichlet, but the proof that $s_ny_{n+1}$ differs depending on if we get our convergence from $(y_n)\to0$ and $(s_n)$ bounded, or $(s_n) \to s$ and $(y_n)$ bounded.

    The proof is almost identical to Abel's test, bound $|s_n|<M$ and use the triangle inequality on the right hand side to get (note $y_k - y_{k+1} > 0$ because decreasing)
    $$
    \sum_{k=1}^n |s_k|(y_k - y_{k+1}) \le M \sum_{k=1}^n (y_k - y_{k+1}) = M(y_1 - y_{n+1}) \le My_1
    $$
    Thus $\sum s_k(y_k - y_{k+1})$ is bounded and increasing, so it converges by MCT.
    To see the $s_ny_{n+1}$ term converges, simply note that $(y_n) \to 0$ and $|s_n|<M$.

    Thus the original series converges, and furthermore
    $$
    \sum_{k=1}^\infty x_ky_k  = \sum_{k=1}^\infty s_k(y_k - y_{k+1})
    $$
    Because $(s_ny_{n+1}) \to 0$.
  \item
    Let $a_n \ge 0$ with $a_1 \ge a_2 \ge \dots \ge 0$ and $\lim a_n = 0$.
    The series $\sum (-1)^n$ is bounded, so Dirichlet's test implies $\sum (-1)^n a_n$ converges.
  }
\end{solution}
